% !TEX root = relative/or/absolute/path/to/root/file.tex
\documentclass[a4paper,12pt]{report}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[utf8]{inputenc}
\usepackage[a4paper, margin=1in]{geometry}

\usepackage{setspace}   % For line spacing
\usepackage{parskip}    % For better paragraph spacing
\usepackage{titlesec}   % For customizing section titles



\title{Chess Engine\\
{\Large COMP4121}}
\author{Joshua Jooeon Shim \\ z5479929}
\date{November 2024}

\begin{document}
\maketitle
\chapter*{Abstract}

Chess is a simple game that is played on a $8 \times 8$ grid board, with 6 different types of pieces that each move differently. Intuitively it is a simple game. However due to the permutational complexity of this game, a large number of humans have dedicated their lives to chess and yet chess remains `unsolved'. Chess remains a highly theoretical board game where it remains subjective in many cases as to which move is the best.
\\\\
By the late 1980's computer chess programs had begun being competitive with top human players and 25 years later, the last known win against a top-performing computer by a human player was recorded in the Ponomariov vs Fritz game.
\\\\
Since then, as hardware and chess algorithmic theory have had noticeable improvements, the delta between human chess and computer chess programs have only increased.
\\\\
For the COMP4121 major project, I have chosen to make a chess engine from scratch. Coming from a pure computer science background with limited mathematical knowledge, I believe that the creation of a chess engine would allow me to actively learn about both the deterministic and randomised algorithms that allow chess engines to overcome the human brain. Hence I came up with the goal of creating a chess engine that would be able to beat the average human player.
\\\\
Initially, I had planned to create a chess engine that primarily utilised the \textit{Monte Carlo Tree Search} algorithm to search for the best moves. However due to the time constraints and the inability for me to feasibly tune the \textit{Monte Carlo Tree Search} using deep learning meant that a \textit{MCTS} chess engine would be significantly less impressive than a more standard chess engine that utilises the \textit{Min-Max Algorithm}.\newpage
The chess bot that I have created is named \textbf{COMP4121bot} and uses a \textit{Universal Chess Interface (UCI)} interface to communicate with the front-end. I have a Raspberry Pi 3 model B setup that runs the chess engine so that it exists as a bot on \textit{lichess.org} where it is available for play. I will keep it active on the website until 31/12/2024 and highly encourage you to play a game against it. Note that as it is running on lower end hardware, its performance is significantly lower than if it were to run on my desktop. The following game setup works but others may / may not be accepted by my bot:\\
Variant: Standard\\
Time Control: Real time\\
Minutes per side: 10\\
Increment in seconds: 3\\
Game Mode: Casual\\\\
% TODO: create a list of available game modes.
Play on this link (you may need to create an account):\\
https://lichess.org/@/COMP4121bot

\tableofcontents
\include{ch1-board_representation}
\include{ch2-move_generation}
\include{ch3-move_evaluation}
\include{ch4-mcts}
\include{ch5-min_max}
\include{ch6-improvements}
\include{ch7-known_bugs}

\end{document}